---
title: "Introduction to Probability - Distributions"
output:
  html_document:
    df_print: paged
---

Uniform Distribution

```{r}
# Define the parameters
max_trials <- 10000
min_value <- 1
max_value <- 6

# Create a matrix that has max_trials rows and 4 columns. This represents the dice rolls.
random_values <- sample(min_value:max_value, max_trials, replace=T)

# Display the final results
barplot(table(random_values))
```

Binomial Distribution

```{r}
# Define the parameters
max_trials <- 10000
n <- 100
p <- 0.5

k <- 0:n

# Simulated distribution
random_values <- rbinom(max_trials, n, p)

# Exact distribution
b <- dbinom(k, n, p)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,n))
lines(k, b, col='blue')
```

Geometric Distribution

```{r}
# Define the parameters
max_trials <- 10000
n <- 20
p <- 0.2

q <- 1-p
k <- 0:n

# Simulated distribution
random_values <- rgeom(max_trials, p)

# Exact distribution
b <- dgeom(k, p)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,n))
lines(k, b, col='blue')
```

Negative Binomial Distribution

Note: The book counts the total number of trials whereas R counts the number of trials beyond the minimum. The first set of plots uses the native R definition and the second set of plots have been shifted to match the book.

```{r}
# Define the parameters
max_trials <- 10000
n <- 20
p <- 0.25
k <- 2

q <- 1-p
x <- 0:n

# Simulated distribution
random_values <- rnbinom(max_trials, k, p)

# Exact distribution
b <- dnbinom(x, k, p)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,n))
lines(x, b, xlim=c(0,n), col='blue')

plot(table(random_values+k)/max_trials, xlim=c(0,n))
lines(x+k, b, xlim=c(0,n), col='blue')

```

Poisson Distribution

The Poisson distribution can be used as an approximation of the binomial distribution under certain conditions.

```{r}
# Define the parameters
max_trials <- 100000
n <- 100
p <- 0.001
max_x <- 10


lambda <- n*p
k <- 0:max_x

# Simulated distribution
random_values <- rbinom(max_trials, n, p)

# Exact distribution
poisson <- dpois(k, lambda)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,max_x))
lines(k, poisson, col='red')
```
