---
title: "Introduction to Probability - Distributions"
output:
  html_document:
    df_print: paged
---

Uniform Distribution

```{r}
# Define the parameters
max_trials <- 10000
min_value <- 1
max_value <- 6

# Create a matrix that has max_trials rows and 4 columns. This represents the dice rolls.
random_values <- sample(min_value:max_value, max_trials, replace=T)

# Display the final results
barplot(table(random_values))
```

Binomial Distribution

```{r}
# Define the parameters
max_trials <- 10000
n <- 100
p <- 0.5

k <- 0:n

# Simulated distribution
random_values <- rbinom(max_trials, n, p)

# Exact distribution
b <- dbinom(k, n, p)
cmf <- cumsum(b)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,n))
lines(k, b, col='blue')

plot(k, cmf, type='l')
```

Geometric Distribution

```{r}
# Define the parameters
max_trials <- 10000
n <- 20
p <- 0.2

q <- 1-p
k <- 0:n

# Simulated distribution
random_values <- rgeom(max_trials, p)

# Exact distribution
b <- dgeom(k, p)
cmf <- cumsum(b)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,n))
lines(k, b, col='blue')

plot(k, cmf, type='l')
```

Negative Binomial Distribution

Note: The book counts the total number of trials whereas R counts the number of trials beyond the minimum. The first set of plots uses the native R definition and the second set of plots have been shifted to match the book.

```{r}
# Define the parameters
max_trials <- 10000
n <- 20
p <- 0.25
k <- 2

q <- 1-p
x <- 0:n

# Simulated distribution
random_values <- rnbinom(max_trials, k, p)

# Exact distribution
b <- dnbinom(x, k, p)
cmf <- cumsum(b)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,n))
lines(x, b, xlim=c(0,n), col='blue')

plot(table(random_values+k)/max_trials, xlim=c(0,n))
lines(x+k, b, xlim=c(0,n), col='blue')

plot(x, cmf, type='l')

plot(x+k, cmf, type='l',xlim=c(0,n))

```

Poisson Distribution

The Poisson distribution can be used as an approximation of the binomial distribution under certain conditions.

```{r}
# Define the parameters
max_trials <- 100000
n <- 100
p <- 0.001
max_x <- 10


lambda <- n*p
k <- 0:max_x

# Simulated distribution
random_values <- rbinom(max_trials, n, p)

# Exact distribution
poisson <- dpois(k, lambda)
cmf <- cumsum(poisson)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,max_x))
lines(k, poisson, col='red')

plot(k, cmf, type='l', xlim=c(0,max_x))

```

Hypergeometric

Book's notation: $h(N, k, n, x)$ where $N$ is the total number of balls, $k$ is the number of blue balls, $n$ is the number of drawn balls, and $x$ is the number of drawn balls that are blue.

R's notation: dhyper(x, m, n, k) where m is the number of blue balls, n is the number of red balls,  k is the number of drawn balls, and x is the number of drawn balls that are blue. 

```{r}
# Define the parameters
max_trials <- 100000

# Setting up with the book's notation
N <- 100
k <- 30
n <- 20

# Translating to R's notation:
m.R <- k
n.R <- N-k
k.R <- n

x <- 0:n

# Simulated distribution
random_values <- rhyper(max_trials, m.R, n.R, k.R)

# Exact distribution
hyper <- dhyper(x, m.R, n.R, k.R)
cmf <- cumsum(hyper)

# Display the final results

plot(table(random_values)/max_trials, xlim=c(0,n))
lines(x, hyper, col='red')

plot(x, cmf, type='l', xlim=c(0,n))

```








