---
title: "Graphing Distributions in R"
output:
  html_document:
    df_print: paged
---

### Frequency Table: Single values only

The plyr library gives us access to a function that counts the results.

```{r}
# Set Parameters
require(plyr)
max_trials <- 100

# Generate Data
possible_outcomes <- c('A', 'B', 'C')
results <- sample(possible_outcomes, max_trials, replace=T)

# Show Frequency Table
table(results)
count(results)
```

### Frequency Table: Multiple values from dataframe

When we have a dataframe, sometimes we want to count the number of times particular combinations arise. This code creates a data frame with random data and then shows different ways to count it.

```{r}
require(plyr)

# Set parameters
max_rows <- 100

# Generate Data
possible_outcomes_1 <- c('A', 'B', 'C')
possible_outcomes_2 <- c('X', 'Y', 'Z')
col1 <- sample(possible_outcomes_1, max_rows, replace=T)
col2 <- sample(possible_outcomes_2, max_rows, replace=T)
df <- data.frame('col1' = col1, 'col2' = col2)

df

# Show Frequency Tables
count(df, 'col1')
count(df, 'col2')
count(df, c('col1', 'col2'))
```

### Pie Chart

Pie charts are created automatically as long as you have the quantities. If you want to do label the percentages, you will need to do that manually.

```{r}
# Set Parameters
num_wedges <- 7
min_val <- 20
max_val <- 40

# Generate random data
quantities <- runif(num_wedges, min = min_val, max = max_val)

# Create labels
section_labels <- paste(rep('Label', num_wedges), 1:num_wedges)

# Create dynamically calculated percentages
percentages <- round(quantities/sum(quantities)*100, 1)
graph_labels <- paste(section_labels, ': ', percentages, '%', sep='')

# Draw pie chart
pie(quantities, labels = graph_labels, main = 'Pie Chart')


```

### Bar Chart: Numbers

```{r}
# Set Parameters
max_trials <- 100

# Generate Data
possible_outcomes <- c('A', 'B', 'C')
results <- sample(possible_outcomes, max_trials, replace=T)

# Show Frequency Table
barplot(table(results), main='Frequency Table')
barplot(table(results), main='Horizontal Frequency Table', horiz=T)
```

### Stem and Leaf Chart:

R's stem and leaf plots sometimes skip values in the stem. You need to recognize the missing values when the digits on the leaves cycle.

```{r}

# Set Parameters
num_trials <- 30
min_val <- 20
max_val <- 40

# Generate random data
quantities <- runif(num_trials, min = min_val, max = max_val)

# Show stem and leaf chart
stem(quantities)
```

### Histograms:

```{r}
# Set Parameters
num_trials <- 1000
mu <- 0
sigma <- 1

# Generate random data
results <- rnorm(num_trials, mean=mu, sd=sigma)

# Show Historgram
hist(results)
hist(results, breaks=seq(from=-5, to=5, by=0.5))

dens <- density(results)
plot(dens$x, length(results)*dens$y, type='l')

```

### Frequency Polygon:

If you are just plotting a function, this is just like plots from earlier. A density plot is given under histogram. It's a bit much to explain.

```{r}
# Set Parameters
num_trials <- 1000
mu <- 0
sigma <- 1

# Generate random data
results <- rnorm(num_trials, mean=mu, sd=sigma)

# Show Historgram
hist(results)
hist(results, breaks=seq(from=-5, to=5, by=0.5))

```


### Box Plots:

A boxplot is best through a data frame when you have multiple columns of data.


```{r}
# Set Parameters
num_trials <- 10

gender <- sample(c('M','F'), num_trials, replace=T)
number <- sample(20:50, num_trials, replace=T)

df <- data.frame('Gender' = gender, 'Number' = number)

boxplot(Number~Gender, data=df)

```

